---
title: "Mathematical Methods for International Commerce "
subtitle: "<br/>Week 12/1: Indefinite Integration (6.1)"
author: "Iegor Vyshnevskyi, Ph.D."
institute: "Sogang University"
date: "May 21, 2025"
output:
  xaringan::moon_reader:
    #css: [default, metropolis, metropolis-fonts]
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

<style>
@media print{
  body, html, .remark-slides-area, .remark-notes-area {
    height: 100% !important;
    width: 100% !important;
    overflow: visible;
    display: inline-block;
    }
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 34px;
    padding: 1em 4em 1em 4em;
}
</style>

<style type="text/css">
.my-one-page-font {
  font-size: 28px;
}
</style>

</style>

<style type="text/css">
.my-one-page-font-table {
  font-size: 24px;
}
</style>


```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

opts_chunk$set(fig.width = 10, 
               message = FALSE, 
               warning = FALSE,
               echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
#install.packages("xaringanthemer")
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "500", "550i"),
  code_font_google   = google_font("Fira Mono"),
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
)
)
```

# Why It Matters in Economics & Finance
- Integration helps in determining total cost, total revenue, and accumulated savings over time.

- It is crucial in calculating areas under curves, which represent total accumulated quantities.

- Applications include:
  - Finding total cost from marginal cost
  - Determining total revenue from marginal revenue
  - Estimating accumulated savings from marginal propensity to save

---

# Learning Objectives
- Recognize the notation for indefinite integration

- Integrate power functions and exponential functions

- Apply integration to find total cost and total revenue

- Apply integration to determine consumption and savings functions

- Use inspection for integrating complex functions

---

# Understanding Indefinite Integration
- Indefinite integration is the reverse of differentiation.

- The general form:
  $$ \int f(x) \, dx = F(x) + C $$
  where:
  - $f(x)$ is the integrand
  - $F(x)$ is the antiderivative
  - $C$ is the constant of integration

<span style="font-size: 60%;">**Note:** The antiderivative, also known as the indefinite integral, is a function whose derivative equals the original function $f(x)$. In essence, finding the antiderivative reverses the process of differentiation.</span>

---

# Understanding Indefinite Integration (cont.)

- Example:

  - If $f(x) = 3x^2$, then:

  $$
  \int 3x^2 \, dx = x^3 + C
  $$

  - Why? Because $\frac{d}{dx}(x^3 + C) = 3x^2$.

---

# The Constant of Integration

- Why do we add $+C$?
  - Represents an **arbitrary constant** since differentiation eliminates constants.

- Example:

  - $\int 3x^2 \, dx = x^3 + C$
  - Both $x^3 + 5$ and $x^3 - 7$ differentiate to $3x^2$.

- From economics perspective, $C$ can represent fixed costs, initial revenue/values, or other baseline values in economic models.
- In practice, we often determine $C$ using initial conditions or boundary values.

---

# Power Rule for Indefinite Integrals

- General formula:

$$
\int x^n \, dx = \frac{x^{n+1}}{n+1} + C \quad (n \neq -1)
$$

- Examples:

  - $\int x^5 \, dx = \frac{x^6}{6} + C$

  - $\int x^{-3} \, dx = \frac{x^{-2}}{-2} + C = -\frac{1}{2x^2} + C$

---

# Integration of Exponential Functions

- Formula:

$$
\int e^{mx} \, dx = \frac{1}{m} e^{mx} + C
$$

- Example:

  - $\int e^{2x} \, dx = \frac{1}{2} e^{2x} + C$

  - $\int e^{-0.5x} \, dx = -2 e^{-0.5x} + C$

---

# Practical Example: Marginal Revenue to Total Revenue (1)

- Suppose the **marginal revenue function** is given by:

$$
MR(x) = 50 - 2x
$$

- To find the total revenue function:

$$
TR(x) = \int (50 - 2x) \, dx
$$

$$
= 50x - x^2 + C
$$

- **Economic Interpretation:**
  - The area under the MR curve represents the total revenue.

---

# Practical Example: Marginal Revenue to Total Revenue (2)
- If the marginal revenue is given by:
  $$ MR(x) = 12 - 2x $$

- The total revenue function is:
  $$ TR(x) = \int (12 - 2x) \, dx = 12x - x^2 + C $$

- Interpretation:
  - The total revenue function represents the total income generated from selling $x$ units of a good.
  - The constant $C$ can be interpreted as fixed costs or initial revenue.
  - The area under the marginal revenue curve gives the total revenue.
  - The total revenue function changes quadratically with respect to $x$.

---

# Practical Example: Marginal Cost to Total Cost (1)

- If the marginal cost is given by:
  $$ MC(x) = 3x^2 + 5 $$

- The total cost function is found by integrating:
  $$ TC(x) = \int (3x^2 + 5) \, dx = x^3 + 5x + C $$

---

# Practical Example: Marginal Cost to Total Cost (2)

- Given the marginal cost function:

$$
MC(x) = 4x^2 - 3x + 7
$$

- Integrate to find the total cost function:

$$
$$ TC(x) = \int (4x^2 - 3x + 7) \, dx = \frac{4x^3}{3} - \frac{3x^2}{2} + 7x + C
$$

---

# Economic Example: Consumption & Savings
- Given the marginal propensity to consume (MPC):

$$
MPC = 0.8
$$

- Integrate to find the consumption function:

$$
\int 0.8 \, dY = 0.8Y + C
$$

- The savings function can be derived similarly by considering the marginal propensity to save (MPS).

---

# Method of Inspection
- For more complex integrals, identify a form that simplifies to a known pattern:

$$
\int (5x^2 + 3x + 2) \, dx = \int 5x^2 \, dx + \int 3x \, dx + \int 2 \, dx
$$

$$
= \frac{5x^3}{3} + \frac{3x^2}{2} + 2x + C
$$
---

# Practice Problems
1. Integrate:
   - (a) $\int 4x^3 \, dx$
   - (b) $\int e^{2x} \, dx$

2. Find the total cost given $MC = 5x^2 - 3x + 4$.

3. Determine the total revenue function given $MR = 15 - 0.5x$.

4. Given the marginal propensity to consume is $MPC = 0.8$, determine the consumption function if initial consumption is 100.

???

# Solutions to Practice Problems

## Problem 1: Integrate the Following

### (a) $\int 4x^3 \, dx$

- Apply the power rule:
$$
\int 4x^3 \, dx = rac{4x^{3+1}}{3+1} + C = rac{4x^4}{4} + C = x^4 + C
$$

**Answer:** $x^4 + C$

#---

### (b) $\int 2e^{0.1x} \, dx$

- Apply the exponential rule:
$$
\int 2e^{0.1x} \, dx = rac{2}{0.1} e^{0.1x} + C = 20 e^{0.1x} + C
$$

**Answer:** $20 e^{0.1x} + C$

#---

# Problem 2: Marginal Cost to Total Cost

- Given $MC(x) = 20 + 4x$

To find the total cost function $C(x)$:

$$
C(x) = \int (20 + 4x) \, dx
$$

- Apply linearity of integration:
$$
= 20x + rac{4x^2}{2} + C
$$

$$
= 20x + 2x^2 + C
$$

**Answer:** $C(x) = 20x + 2x^2 + C$

#---

# Problem 3: Marginal Propensity to Consume

- Given $MPC = 0.8$ and initial consumption $C_0 = 100$

The consumption function $C(Y)$ can be determined by integrating the MPC:

$$
C(Y) = \int 0.8 \, dY = 0.8Y + C
$$

- Since initial consumption is 100 when $Y = 0$:
$$
100 = 0.8(0) + C \implies C = 100
$$

Thus, the consumption function is:

$$
C(Y) = 0.8Y + 100
$$

**Answer:** $C(Y) = 0.8Y + 100$

#---


---

# Summary
- Indefinite integration helps in finding total values given marginal functions.

- Applications in economics include calculating total cost, total revenue, and accumulated savings.

- Key formulas:
  - Power Rule
  - Exponential Function
  - Method of Inspection

---

class: inverse, center, middle

# 2. Group Activity

---


# Math Battle: Indefinite Integrals

**Objective:** Apply indefinite integration concepts to solve economic problems.  
**Format:** 4 groups, 4 students each. Each group receives a distinct problem.  
**Time:** 20 minutes to solve, 5 minutes to present.

### Math Battle Rules:

- Each group presents their solutions. 
- Other groups can challenge the solution if they find discrepancies.
- Correct challenges earn 2 points; incorrect challenges lose 1 point.
- Winning group gets a prize!

---

# Group 1: Marginal Cost Analysis

A firm’s marginal cost function is given by:

$$ MC(x) = 15x^2 - 40 $$

1. Determine the total cost function by integrating the MC function.

2. Identify the constant of integration given that the fixed cost is $120.

---

# Group 2: Consumption and Savings

The marginal propensity to save is given by:

$$ MPS = 0.2x - 0.05 $$

1. Integrate the MPS to find the savings function.

2. If the initial savings is $50, determine the complete savings function.

---

# Group 3: Investment Accumulation

A firm invests $200 annually in a project, and the marginal investment function is given by:

$$ MI(t) = 200 e^{0.03t} $$

1. Integrate to find the total investment function.

2. Determine the accumulated investment after 5 years.

---

# Group 4: Demand Function Analysis

A product’s marginal demand function is given by:

$$ MD(Q) = -3Q^2 + 12Q - 5 $$

1. Integrate to find the demand function.

2. Determine the constant of integration if the initial demand is $50.

???



# Group 1: Cost Analysis - Solutions

### Problem:
Integrate the marginal cost function to find the total cost function.

**Given:** \( MC(x) = 12x^2 - 5x + 3 \)

**Solution:**

\[\int (12x^2 - 5x + 3) \, dx = \int 12x^2 \, dx - \int 5x \, dx + \int 3 \, dx\]

\[\int 12x^2 \, dx = 12 \cdot rac{x^3}{3} = 4x^3\]

\[\int -5x \, dx = -5 \cdot rac{x^2}{2} = -rac{5x^2}{2}\]

\[\int 3 \, dx = 3x\]

**Total Cost Function:**

\[	ext{TC}(x) = 4x^3 - rac{5x^2}{2} + 3x + C\]

#---

# Group 2: Consumption Analysis - Solutions

### Problem:
Find the consumption function given the MPC and initial consumption.

**Given:** \( MPC = 0.75 \), Initial consumption = 120

**Solution:**

Marginal Propensity to Consume:

\[\int MPC \, dY = \int 0.75 \, dY\]

\[	ext{C}(Y) = 0.75Y + C\]

**Finding the Constant \( C \):**

At \( Y = 0 \), \( C = 120 \).

**Consumption Function:**

\[	ext{C}(Y) = 0.75Y + 120\]

#---

# Group 3: Revenue Analysis - Solutions

### Problem:
Integrate the marginal revenue function to find the total revenue function.

**Given:** \( MR(x) = 20x - 3 \)

**Solution:**

\[\int (20x - 3) \, dx = \int 20x \, dx - \int 3 \, dx\]

\[\int 20x \, dx = 20 \cdot rac{x^2}{2} = 10x^2\]

\[\int -3 \, dx = -3x\]

**Total Revenue Function:**

\[	ext{TR}(x) = 10x^2 - 3x + C\]

#---

# Group 4: Demand Analysis - Solutions

### Problem:
Find the demand function given the marginal demand function.

**Given:** \( MD(Q) = -4Q + 15 \)

**Solution:**

\[\int (-4Q + 15) \, dQ = \int -4Q \, dQ + \int 15 \, dQ\]

\[\int -4Q \, dQ = -4 \cdot rac{Q^2}{2} = -2Q^2\]

\[\int 15 \, dQ = 15Q\]

**Demand Function:**

\[	ext{D}(Q) = -2Q^2 + 15Q + C\]

#---


---



class: inverse, center, middle

# 3. Home work #2

---

class: my-one-page-font-table

## Homework #2

- **Due Date:** June 13, 2025, before the start of class.
- **Submission Format:** Submit your solutions as a single PDF file via the Cyber Campus.
- **Instructions:** 
  - Clearly show all steps and calculations.
  - Include explanations for your answers where applicable.
  - Ensure your submission is neat and well-organized.
  - Bring any questions to the office hours or email me.

- Work on your Home Assignment #2 (Jacques, 10th edition, Chapters 5-9):
  - Chapter 5.1: Exercise 5.1, Problem 5 (p. 411)
  - Chapter 5.2: Exercise 5.2, Problem 7 (p. 427)
  - Chapter 5.3: Exercise 5.3, Practice Problem inside the chapter (p. 432 only)
  - Chapter 5.4: Exercise 5.4, Problem 5 (p. 457)
  - Chapter 5.5: Exercise 5.5, Problems 6 and 7  (p. 467)
  - Chapter 5.6: Exercise 5.6, Problem 4 (p. 479)
  - Chapter 6.1: Exercise 6.1, Problem 3 (p. 506)
  - Chapter 6.2: Exercise 6.2, Problem 6 (p. 521)
  - Chapter 7.1: Exercise 7.1, Problem 5 (p. 553)
  - Chapter 7.2: Exercise 7.2, Problem 6 (p. 572)
  - Chapter 7.3: Exercise 7.3, Problem 5 (p. 584)
  - Chapter 8.1: Exercise 8.1, Problem 6 (p. 615)
  - Chapter 8.2: Exercise 8.2, Problem 3 (p. 626)
  - Chapter 9.1: Exercise 9.1, Problem 4 (p. 655)
  - Chapter 9.2: Exercise 9.2, Problem 3 (p. 670)

Good luck!


---

class: inverse, center, middle

# Any QUESTIONS?

## Thank you for your attention!

---

## Next Classes

- (May 23) Definite Integration (6.2) 

## Reminder:

- **Quiz 2**: May 28, 2025



???
1. To print pdf slides
https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

pagedown::chrome_print("W1_ME.html") # but not all pictures are visible

2. Option: https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

install.packages("remotes")
remotes::install_github("jhelvy/xaringanBuilder")
remotes::install_github("jhelvy/renderthis@v0.0.9")

library(xaringanBuilder)
build_pdf("DVC.html")

3. Option
writeBin(as.raw(c()), "favicon.ico") # create an empty favicon.ico file
install.packages("renderthis")
remotes::install_github('rstudio/chromote')
library(renderthis)

renderthis::to_pdf("W12_1_ME.html")

getwd()
setwd("C:/Users/Iegor/OneDrive - kdis.ac.kr/Documents/GitHub/Sogang/2025/Spring/Mathematical Economics/Week 12_1")
