---
title: "Mathematical Methods for International Commerce "
subtitle: "<br/>Week 4/2:  Geometric Series and Investment Appraisal"
author: "Iegor Vyshnevskyi, Ph.D."
institute: "Sogang University"
date: "March 28, 2025"
output:
  xaringan::moon_reader:
    #css: [default, metropolis, metropolis-fonts]
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

<style>
@media print{
  body, html, .remark-slides-area, .remark-notes-area {
    height: 100% !important;
    width: 100% !important;
    overflow: visible;
    display: inline-block;
    }
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 34px;
    padding: 1em 4em 1em 4em;
}
</style>

<style type="text/css">
.my-one-page-font {
  font-size: 28px;
}
</style>

</style>

<style type="text/css">
.my-one-page-font-table {
  font-size: 24px;
}
</style>


```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

opts_chunk$set(fig.width = 10, 
               message = FALSE, 
               warning = FALSE,
               echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
#install.packages("xaringanthemer")
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "500", "550i"),
  code_font_google   = google_font("Fira Mono"),
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
)
)
```

# Why This Topic Matters

Understanding **geometric series** and **investment appraisal** is crucial in economics and finance:

- **Geometric series** model regular payments, compounding, and amortization schedules

- **Investment appraisal** tools help evaluate project viability, return potential, and loan value

- Core to decisions in **banking, government budgeting, business strategy, and personal finance**

- Equips you to analyze time-based value of money, compare alternatives, and manage risk

> “A dollar today is worth more than a dollar tomorrow — let’s understand why.”

---

class: inverse, center, middle

# Section 3.3: Geometric Series

---

## What is a Geometric Progression?

A **geometric sequence** has the form:

$$
a, ar, ar^2, ar^3, \dots
$$

Where:
- $a$ = first term
- $r$ = common ratio (can be fraction or negative)

Basically, each term is the previous term multiplied by the common ratio.

## Example: $2, 4, 8, 16, 32$

$a = 2, r = 2$

---

## Sum of a Finite Geometric Series

$$
S_n = a \cdot \frac{1 - r^n}{1 - r}, \quad r \ne 1
$$

- Example: $5 + 10 + 20 + 40$  
$a = 5, r = 2, n = 4$

$$
S_4 = 5 \cdot \frac{1 - 2^4}{1 - 2} = 5 \cdot \frac{-15}{-1} = 75
$$

---

## Example: Regular Savings Plan

Annual deposit: $100, interest = 5%, duration = 5 years

It means that each year, you save $100 and earn 5% interest.

$$
S_5 = 100 \cdot \frac{1.05^5 - 1}{0.05} \approx 100 \cdot 5.526 = \boxed{\$552.60}
$$

In 5 years, you will have saved $552.60.

---

## Your turn! Calculate the Sum

- Find the sum of the following geometric series:

$$
2 + 4 + 8 + 16 + 32
$$

$$
3 + 6 + 12 + 24 + 48
$$

- Find the final Savings Plan amount for $200/year, 4% interest, 10 years

---

## Example: Loan Repayment by Instalments

Loan = $2000, repaid in 4 annual instalments at 10%

$$
x \cdot \left( \frac{1 - (1.10)^{-4}}{0.10} \right) = 2000
$$

Solve for $x$ to find equal annual instalments.

$$
x = 2000 \cdot \frac{0.4641}{0.10} \approx 928.20
$$

---

## Your turn! Calculate the Instalments

- A loan of $5000 is repaid in 5 annual instalments at 8%. Find the annual instalment.


- A loan of $3000 is repaid in 3 annual instalments at 6%. Find the annual instalment.

---

class: inverse, center, middle

# Section 3.4: Investment Appraisal

---

## Investment Appraisal

Stands for evaluating the financial viability of investments.

## Present Value (Discrete)

- Present value $PV$ is the current value of future cash flows.

$$
PV = \frac{A}{(1 + r)^n}
$$

- $1000$ in 3 years at 6%:

$$
PV = \frac{1000}{1.06^3} \approx 839.62
$$

---

## Present Value (Continuous)

$$
PV = A \cdot e^{-rt}
$$

- $1000$ in 3 years at 6%:

$$
PV = 1000 \cdot e^{-0.18} \approx 836.00
$$

### What's the difference between discrete and continuous?

- **Discrete**: Values occur at specific intervals (e.g., yearly, monthly).  
  Example: Annual interest compounding.

- **Continuous**: Values change smoothly over time, modeled using exponential functions.  
  Example: Continuous compounding of interest.

---

## Net Present Value (NPV)

- Net present value $NPV$ is the sum of present values of cash flows.

$NPV = -C_0 + \sum_{t=1}^n \frac{C_t}{(1 + r)^t}$

- Example: Cost = $2000, Returns = $800/year for 3 years, $r = 10\%$:

$$
NPV \approx -2000 + \frac{800}{1.10} + \frac{800}{1.10^2} + \frac{800}{1.10^3} \approx 75.13
$$

---

## Present Value of an Annuity

- Annuity is a series of equal payments over time.

$$
PV = PMT \cdot \frac{1 - (1 + r)^{-n}}{r}
$$

- $1000/year for 5 years at 6%:

$$
PV = 1000 \cdot \frac{1 - (1.06)^{-5}}{0.06} \approx 4212.40
$$

---

## Internal Rate of Return (IRR)

- IRR is the rate at which NPV = 0.

- We calculate IRR to compare investment options.

Find $r$ such that $NPV = 0$:

$$
0 = -C_0 + \sum \frac{C_t}{(1 + r)^t}
$$

Solve numerically or using software.

---

## Present Value of Government Bonds

Bond pays $50/year for 10 years, plus $1000 at end.

$$
PV = \sum_{t=1}^{10} \frac{50}{(1.05)^t} + \frac{1000}{(1.05)^{10}} \approx 925.81
$$

---

# Practice Problems

1. Sum of: $3 + 6 + 12 + 24$

2. $200/quarter, 4% quarterly compounding for 3 years

3. PV of $1500 in 2 years at 8% (continuous)

4. NPV: $3000 cost, $1200/year for 3 years, 7% rate

5. PV of $500 annuity for 6 years at 5%

6. IRR: $2000 cost, $500/year for 5 years

---

# Summary

- Geometric series = finance modeling

- Investment tools = PV, NPV, IRR

- Practical for decision-making in public/private sectors

---

class: inverse, center, middle

# Any QUESTIONS?

## Thank you for your attention!

---

# Next Class

- (April 2) The Derivative of Functions (4.1), Rules of Differentiation (4.2) 

# Next Friday (April 4) - Quiz #1

- Please review your Home Work #1, in-class problems and problems from the textbook. 


???
1. To print pdf slides
https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

pagedown::chrome_print("W1_ME.html") # but not all pictures are visible

2. Option: https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

install.packages("remotes")
remotes::install_github("jhelvy/xaringanBuilder")
remotes::install_github("jhelvy/renderthis@v0.0.9")

library(xaringanBuilder)
build_pdf("DVC.html")

3. Option
writeBin(as.raw(c()), "favicon.ico") # create an empty favicon.ico file
install.packages("renderthis")
remotes::install_github('rstudio/chromote')
library(renderthis)

renderthis::to_pdf("W4_2_ME.html")

getwd()
setwd("C:/Users/Iegor/OneDrive - kdis.ac.kr/Documents/GitHub/Sogang/2025/Spring/Mathematical Economics/Week 4_2")
