---
title: "Mathematical Methods for International Commerce "
subtitle: "<br/>Week 7/2: The Derivative of the Exponential and Natural Log Functions (4.8)"
author: "Iegor Vyshnevskyi, Ph.D."
institute: "Sogang University"
date: "April 18, 2025"
output:
  xaringan::moon_reader:
    #css: [default, metropolis, metropolis-fonts]
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

<style>
@media print{
  body, html, .remark-slides-area, .remark-notes-area {
    height: 100% !important;
    width: 100% !important;
    overflow: visible;
    display: inline-block;
    }
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 34px;
    padding: 1em 4em 1em 4em;
}
</style>

<style type="text/css">
.my-one-page-font {
  font-size: 28px;
}
</style>

</style>

<style type="text/css">
.my-one-page-font-table {
  font-size: 24px;
}
</style>


```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

opts_chunk$set(fig.width = 10, 
               message = FALSE, 
               warning = FALSE,
               echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
#install.packages("xaringanthemer")
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "500", "550i"),
  code_font_google   = google_font("Fira Mono"),
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
)
)
```

# Agenda  

1. The Derivative of the Exponential and Natural Log Functions (4.8) 

2. Individual Activity: Solve the Problems!

3. Midterm Exam Review

---

class: inverse, center, middle

# 1. The Derivative of the Exponential and Natural Log Functions (4.8)

---

# Why It Matters: Economics & Finance Perspective

Exponential and logarithmic functions are foundational in economics and finance:

- They model **continuous growth and decay** – essential for understanding investment returns, population growth, inflation, and depreciation.
- The exponential function captures **compound interest**, while logarithms are key for solving equations involving **growth rates**.
- Derivatives of these functions help economists and analysts assess **rates of change** – such as marginal returns, discounting future values, or analyzing production over time.

Understanding how to **differentiate** exponential and logarithmic functions equips you with tools to interpret trends, optimize financial decisions, and make precise economic forecasts.

---

# Learning Objectives

## Section 4.8 – Derivatives of Exponential and Logarithmic Functions

- Differentiate the **exponential function**
- Differentiate the **natural logarithm function**
- Use **chain, product and quotient rules** with these functions
- Apply exponential models to **real-world economic problems**

---

# Basic Rules of Differentiation (Exponential & Logarithmic Functions)

### Exponential Function:
The exponential function grows **proportionally to its value**.

$$
\frac{d}{dx}e^x = e^x
$$
This is unique: the derivative of $e^x$ is itself!

### Natural Logarithm:
Logarithms are used to reverse exponential growth.

$$
\frac{d}{dx}\ln(x) = \frac{1}{x}, \quad x > 0
$$

---

# Basic Rules of Differentiation (Exponential & Logarithmic Functions) (cont'd)

### Chain Rule:
Use when differentiating **functions within functions**.

- $$ \frac{d}{dx} e^{f(x)} = f'(x) \cdot e^{f(x)} $$
- $$ \frac{d}{dx} \ln(f(x)) = \frac{f'(x)}{f(x)} $$

These rules are essential in **economic models** involving growth, decay, and elasticity.

---

class: my-one-page-font-table

# Examples: Basic Differentiation

### 1. Differentiate:
$$ f(x) = e^{3x} $$
**Solution:**
$$ f'(x) = 3e^{3x} $$
- Use chain rule: derivative of exponent times original exponential

### 2. Differentiate:
$$ g(x) = \ln(5x^2 + 1) $$
**Solution:**
$$ g'(x) = \frac{10x}{5x^2 + 1} $$
- Chain rule: differentiate inside function $5x^2 + 1$ → $10x$

**Tip:** Exponentials often model compounded growth, and logs appear in utility, elasticity, and returns.

---

# Product & Quotient Rule in Action

## Example 1: Product Rule
$$ y = x^2 e^x $$
Apply:
$$ \frac{d}{dx}(uv) = u'v + uv' $$

- $u = x^2$, $v = e^x$
- Derivative: $$ \frac{dy}{dx} = 2x e^x + x^2 e^x = e^x(2x + x^2) $$

Use when both parts involve variables (common in cost/revenue products).

---
# Product & Quotient Rule in Action (cont'd)

## Example 2: Quotient Rule
$$ y = \frac{\ln(x)}{x^2} $$
Apply:
$$ \frac{d}{dx} \left(\frac{u}{v}\right) = \frac{u'v - uv'}{v^2} $$

- $u = \ln(x)$, $v = x^2$
- Derivative:
$$
\frac{dy}{dx} = \frac{1/x \cdot x^2 - \ln(x) \cdot 2x}{x^4} = \frac{x - 2x\ln(x)}{x^4}
$$

Use when differentiating ratios like **marginal utility/cost per unit**.

---

# Application: Continuous Revenue Growth

Let the revenue function be:
$$ R(t) = 5000e^{0.05t} $$
This implies revenue grows **at a continuous rate of 5%** per time unit.

### Find the rate of change of revenue:
$$
R'(t) = 5000 \cdot 0.05e^{0.05t} = 250e^{0.05t}
$$

Interpretation:
- The **growth rate is proportional** to the current revenue.
- Common in modeling investment returns, inflation, or GDP.

---

# Application: Elasticity of Growth

A common growth model:
$$ Q(t) = Ae^{rt} $$
Where:
- $Q(t)$ is output (e.g., capital, population)
- $A$ is initial value
- $r$ is the growth rate

---

# Application: Elasticity of Growth (cont'd)

### Find:
- $$ \frac{dQ}{dt} = rAe^{rt} = rQ(t) $$
- Elasticity of growth:
$$ E = \frac{dQ/dt}{Q} = r $$

**Interpretation:** Elasticity is **constant** in exponential growth → % change in $Q$ for 1% change in time.

Used in modeling:
- Population growth
- Compound interest
- Inflation and real returns

---


# Summary

- Exponential and log functions are common in **growth, interest, decay** models
- Chain rule is crucial when inside other functions
- Product and quotient rules still apply!

> Exponentials capture **compounding**; logs help **linearize** growth patterns.

---

class: inverse, center, middle

# 2. Individual Activity: Solve the Problems! 

---

# Practice Problems

1. Differentiate:
   - (a) $f(x) = e^{2x^2}$
   - (b) $g(x) = \ln(x^2 + 1)$

2. Use product rule:
   - (a) $y = x e^{3x}$

3. Use quotient rule:
   - (a) $y = \frac{e^x}{x^3}$

4. Application:
   - Revenue grows as $R(t) = 12000e^{0.04t}$. Find $R'(t)$ and interpret.

???



# ✅ Solutions: Practice Problems


## 🔹 1. Differentiate:

### (a) $f(x) = e^{2x^2}$

Use chain rule:

$$
f'(x) = e^{2x^2} \cdot \frac{d}{dx}(2x^2) = e^{2x^2} \cdot 4x = 4x e^{2x^2}
$$



### (b) $g(x) = \ln(x^2 + 1)$

Use chain rule:

$$
g'(x) = \frac{1}{x^2 + 1} \cdot 2x = \frac{2x}{x^2 + 1}
$$



## 🔹 2. Product Rule

### (a) $y = x e^{3x}$

Let $u = x$, $v = e^{3x}$

$$
y' = u'v + uv' = 1 \cdot e^{3x} + x \cdot 3e^{3x} = e^{3x} + 3x e^{3x}
$$

✅ Final Answer: $y' = e^{3x}(1 + 3x)$



## 🔹 3. Quotient Rule

### (a) $y = \frac{e^x}{x^3}$

Let $u = e^x$, $v = x^3$

$$
y' = \frac{u'v - uv'}{v^2} = \frac{e^x \cdot x^3 - e^x \cdot 3x^2}{x^6}
= \frac{e^x (x^3 - 3x^2)}{x^6}
= \frac{e^x x^2 (x - 3)}{x^6} = \frac{e^x(x - 3)}{x^4}
$$

✅ Final Answer: $y' = \frac{e^x(x - 3)}{x^4}$



## 🔹 4. Application: Revenue Growth

### Given: $R(t) = 12000e^{0.04t}$

$$
R'(t) = 12000 \cdot 0.04 \cdot e^{0.04t} = 480 e^{0.04t}
$$

### ✅ Interpretation:

The **rate of change of revenue** is proportional to the current revenue level.  
This reflects **exponential growth** – each year, revenue grows faster.


---

class: inverse, center, middle

# 3. Midterm Exam Review 

---

# Midterm Exam Review

## Coverage:
- Basic Algebra & Solving Equations (1.1–1.4)
- Supply and Demand, Transposition, National Income (1.5–1.7)
- Quadratic Functions, Revenue & Profit (2.1–2.2)
- Indices, Logs, Exponentials (2.3–2.4)
- Percentages, Compound Interest (3.1–3.2)
- Geometric Series & Investment Appraisal (3.3–3.4)
- Derivatives & Marginal Functions (4.1–4.3)
- Chain/Product/Quotient Rules, Elasticity (4.4–4.5)
- Optimization (4.6–4.7), Derivatives of Exp/Log (4.8)

**Also Review:** Lecture slides, Homework #1, Quiz #1

---

# What You Should Be Able to Do

- Simplify, factor, solve linear & quadratic equations
- Graph demand, supply, quadratic, exponential functions
- Interpret elasticity, marginal cost/revenue, and APL/MPL
- Apply formulas for compound interest, present value, annuities
- Differentiate power, exponential, log functions
- Use derivative tests to find and classify stationary points

---

# Algebra & Quadratics

- Factor and solve:
  - $x^2 - 5x + 6 = 0$
  - $2x^2 - 3x - 2 = 0$
- Graph: $y = x^2 - 4x + 3$
- Use quadratic formula

---

# Supply & Demand, Income Determination

- Sketch and solve:
  - $Q_d = 100 - 5P$, $Q_s = 20 + 3P$
- Find equilibrium
- National income:
  - $Y = C_0 + cY + I + G$, solve for Y

---

# Logs, Indices, Exponentials

- Simplify:
  - $2^3 \times 2^2$
  - $\ln(e^3)$
- Solve: $3^x = 81 \Rightarrow x = 4$
- Evaluate: $(1.05)^5$

---

# Finance Applications

- Calculate:
  - % increase/decrease
  - Compound value: $A = P(1 + r)^n$
  - NPV: $\text{NPV} = \sum \frac{C_t}{(1 + r)^t} - C_0$
  - Loan instalment (annuity formula)

---

# Derivatives & Optimization

- Power rule: $\frac{d}{dx} x^n$
- Product & chain rule
- Find:
  - Marginal Revenue/Cost
  - When MR = MC
  - Maximize profit using first and second derivative test

---

# Elasticity

- Arc elasticity:
\[
E_d = \frac{\Delta Q / \text{avg } Q}{\Delta P / \text{avg } P}
\]
- Point elasticity:
\[
E_d = \frac{dQ}{dP} \cdot \frac{P}{Q}
\]
- Know how elasticity affects revenue

---

# Practice & Prep

Review your:
  - Lecture slides and solved examples
  - Homework #1
  - Quiz #1

**Follow exam instructions!** No phones allowed. Only one A4 double-sided handwritten cheat sheet (formulas only) is permitted. The time limit will be strictly enforced.

Time to shine – you've got this!

---

class: inverse, center, middle

# Any QUESTIONS?

## Thank you for your attention!

---

## Next Classes

- (April 23) No Class (Midterm Exam Week)
- (April 25) Mid term exam (in class):
  - Review all material from the beginning of the semester
  - Pay attention to the examples in the slides, HW #1, Quiz #1 and the exercises in the textbook


???
1. To print pdf slides
https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

pagedown::chrome_print("W1_ME.html") # but not all pictures are visible

2. Option: https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

install.packages("remotes")
remotes::install_github("jhelvy/xaringanBuilder")
remotes::install_github("jhelvy/renderthis@v0.0.9")

library(xaringanBuilder)
build_pdf("DVC.html")

3. Option
writeBin(as.raw(c()), "favicon.ico") # create an empty favicon.ico file
install.packages("renderthis")
remotes::install_github('rstudio/chromote')
library(renderthis)

renderthis::to_pdf("W7_2_ME.html")

getwd()
setwd("C:/Users/Iegor/OneDrive - kdis.ac.kr/Documents/GitHub/Sogang/2025/Spring/Mathematical Economics/Week 7_2")
