---
title: "Central Banking"
subtitle: "<br/>Week 13: Global Coordination and Central Banking" 
author: "Iegor Vyshnevskyi, Ph.D."
institute: "Sogang University"
date: "May 28, 2025"
output:
  xaringan::moon_reader:
    #css: [default, metropolis, metropolis-fonts]
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
---

<style>
@media print{
  body, html, .remark-slides-area, .remark-notes-area {
    height: 100% !important;
    width: 100% !important;
    overflow: visible;
    display: inline-block;
    }
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 38px;
    padding: 1em 4em 1em 4em;
}
</style>

<style type="text/css">
.my-one-page-font {
  font-size: 32px;
}
</style>

</style>

<style type="text/css">
.my-one-page-font-table {
  font-size: 28px;
}
</style>


```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

opts_chunk$set(fig.width = 10, 
               message = FALSE, 
               warning = FALSE,
               echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
#install.packages("xaringanthemer")
library(xaringanthemer)
style_mono_accent(
  base_color = "#135978", # #1c5253 #135978
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "500", "550i"),
  code_font_google   = google_font("Fira Mono"),
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
)
)
```

# Agenda  

1. Global Coordination and Central Banking

2. Class Activity

---

class: inverse, center, middle

# 1. Global Coordination and Central Banking

---

# Introduction

- Central banks increasingly operate in a globalized financial system, requiring coordination and cooperation.

- This lecture will cover:
  - The evolution of central bank coordination from the Bretton Woods era to the present.
  - The key frameworks and institutions supporting central bank cooperation.
  - The role of the ECB, BIS, IMF, and G20 in international coordination.
  - Real-world examples of cross-border policy alignment.
  - Challenges and implications for monetary policy and financial stability.

---

# Why Global Coordination?

- Mitigating spillover effects from major economies.

- Coordinating monetary policies to prevent financial crises.

- Harmonizing regulatory frameworks to reduce systemic risks.

- Facilitating crisis response through cross-border liquidity support.

- Addressing global financial imbalances through coordinated interventions.

---

# Historical Evolution of Central Bank Coordination

- **Bretton Woods (1944):** Establishment of the IMF and World Bank to stabilize exchange rates and provide financial assistance.

- **Plaza Accord (1985):** Coordinated intervention to devalue the US dollar.

- **Asian Financial Crisis (1997-1998):** IMF-led stabilization packages and coordinated monetary policy.

- **Global Financial Crisis (2008):** Central bank swap lines to provide liquidity.

- **COVID-19 Pandemic (2020):** Joint monetary and fiscal interventions.

- **Current Trends (2025):** Focus on digital currencies, climate risks, trade tariffs and cybersecurity.

---

# Key Institutions and Frameworks

- **BIS ([Bank for International Settlements](https://www.bis.org/about/index.htm?m=2)):** Platform for global monetary and financial [cooperation](https://www.bis.org/topic/green_finance/cooperation.htm).

- **IMF ([International Monetary Fund](https://www.imf.org/en/About/Factsheets/IMF-at-a-Glance)):** Provides financial stability and crisis intervention.

- **Largest economies CBs** (e.g., Fed, ECB, BoJ): Coordinates with G20 and emerging markets.

- **[G20](https://g20.org/about-g20/overview/):** Platform for policy discussions among major economies.

- **FSB ([Financial Stability Board](https://www.fsb.org/about/)):** Monitors and makes recommendations for global financial stability.

---

# Case Study: ECB’s International Cooperation Activities

- Focus on:
  - **G20 Emerging Markets:** Technical assistance and policy advisory.
  - **EU Candidate Countries:** Strengthening central banking frameworks.
  - **Regional Institutions:** Promoting regional financial stability.

- Activities include:
  - Workshops, expert visits, joint research.
  - Staff secondments and training programs.

---

# Case Study: Global Liquidity Swaps (BIS 2022)

- Following the 2008 crisis, major central banks established swap lines to provide USD liquidity.

- Central banks involved:
  - Federal Reserve, ECB, BoJ, Bank of England, SNB.

- Objective: Mitigate USD funding pressures and prevent market disruptions.

---

# Challenges in International Coordination

- **Policy Divergence:** Differing monetary policy goals across regions.

- **Data Sharing:** Data privacy concerns limit cross-border cooperation.

- **Exchange Rate Management:** Conflicts between floating and fixed regimes.

- **Cybersecurity Risks:** Cross-border payment systems vulnerable to attacks.

---

# Current Trends in Central Bank Cooperation

- Rise of **digital currencies and payment systems.**

- Focus on **green finance and climate-related risks.**

- Addressing **cybersecurity risks in interconnected financial networks.**

- Enhancing cross-border payment systems.

---

# Conclusion

- Global coordination is vital to ensuring financial stability in a globalized financial system.

- Institutions like the BIS, IMF, and ECB play key roles in harmonizing monetary policies.

- The rise of digital currencies, climate risks, and cybersecurity challenges will shape future coordination efforts.

- Ongoing challenges include policy divergence, data sharing, and geopolitical tensions.

---


class: inverse, center, middle

# 2. In-class Group Activity

---

class: inverse, center, middle

# Any QUESTIONS?

**Let’s build central banks that serves growth, stability & people.**

## Thank You!  


---

# Next Class

-(June 04) Future Challenges for Central Banks  
      - The readings will be posted on the Cyber Campus website.

# We expect to have a guest speaker from the National bank of Ukraine.


???

1. To print pdf slides
https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

pagedown::chrome_print("W1_ME.html") # but not all pictures are visible

2. Option: https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

install.packages("remotes")
remotes::install_github("jhelvy/xaringanBuilder")
remotes::install_github("jhelvy/renderthis@v0.0.9")

library(xaringanBuilder)
build_pdf("DVC.html")

3. Option
writeBin(as.raw(c()), "favicon.ico") # create an empty favicon.ico file
install.packages("renderthis")
remotes::install_github('rstudio/chromote')
library(renderthis)

renderthis::to_pdf("W13_CB.html")

getwd()
setwd("C:/Users/Iegor/OneDrive - kdis.ac.kr/Documents/GitHub/Sogang/2025/Spring/Central Banking/Week 13")
