---
title: "Financial Institutions and System"
subtitle: "<br/>Week 12: Quantity Theory, Inflation, and the Demand for Money."
author: "Iegor Vyshnevskyi, Ph.D."
institute: "Sogang University"
date: "May 16, 2025"
output:
  xaringan::moon_reader:
    #css: [default, metropolis, metropolis-fonts]
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: '16:9'
mathjax: default
---

<style>
@media print{
  body, html, .remark-slides-area, .remark-notes-area {
    height: 100% !important;
    width: 100% !important;
    overflow: visible;
    display: inline-block;
    }
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 38px;
    padding: 1em 4em 1em 4em;
}
</style>

<style type="text/css">
.my-one-page-font {
  font-size: 32px;
}
</style>

</style>

<style type="text/css">
.my-one-page-font-table {
  font-size: 28px;
}
</style>


```{r setup, include = FALSE}
library(tidyverse)
library(knitr)

opts_chunk$set(fig.width = 10, 
               message = FALSE, 
               warning = FALSE,
               echo = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
#install.packages("xaringanthemer")
library(xaringanthemer)
style_mono_accent(
  base_color = "#135978", # #1c5253
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "500", "550i"),
  code_font_google   = google_font("Fira Mono"),
  colors = c(
  red = "#f34213",
  purple = "#3e2f5b",
  orange = "#ff8811",
  green = "#136f63",
  white = "#FFFFFF"
)
)
```

# Agenda  

1. Quantity Theory, Inflation, and the Demand for Money    

2. Class Activity

---

class: inverse, center, middle

# 1. The Foreign Exchange Market

---

# Introduction

### Quantity Theory, Inflation, and the Demand for Money

- Identify the factors underlying the portfolio choice theory of money demand.

- Assess and interpret the empirical evidence on the validity of the liquidity preference and portfolio theories of money demand.

- Understand the relationship between money supply, price level, and velocity.

.small[Note: recall that liquidity preference theory ( the demand for money is prioritized over other assets. The interest rate in an economy is influenced by the supply of and demand for money.) is a Keynesian approach to money demand, while portfolio theory (investors who make their decisions based solely on expected return (the mean or average return) and volatility (standard deviation) should make rational choices) is a more modern approach.]


---

class: my-one-page-font

# Quantity Theory of Money - Equation of Exchange

- The Quantity Theory of Money connects the money supply (M), velocity (V), price level (P), and output (Y).

**Equation of Exchange:**

$$ M \times V = P \times Y $$

- Velocity (V, measures the number of times that one unit of currency is used to purchase goods and services within a given time period) is assumed constant in the short run.
- Implications:
  - Changes in money supply (M) lead to proportional changes in price level (P) when output (Y) is constant.
  - If output is at full employment, changes in money supply affect price level directly.

- If output is flexible, changes in money supply can affect both price level and output.
- The equation of exchange is a fundamental concept in monetary economics.
- It provides a framework for understanding the relationship between money supply, price level, and output.
- It is used to analyze the effects of monetary policy on inflation and economic activity.
- The equation of exchange is derived from the identity that states that the total amount of money in circulation multiplied by the number of times it is spent equals the total value of goods and services produced in an economy.

---

# Quantity Theory of Money - example

If nominal GDP (PY) is $10 trillion, the money supply (M) is $2 trillion, then:
- Velocity (V) can be calculated as:
$$ V = \frac{PY}{M} = \frac{10 \text{ trillion}}{2 \text{ trillion}} = 5 $$
- This means that each dollar in the money supply is spent 5 times in a year.

- If the money supply (M) increases to $3 trillion, and velocity (V) remains constant, then:
$$ P \times Y = M \times V $$
$$ P \times Y = 3 \text{ trillion} \times 5 $$

- If output (Y) is constant at $10 trillion, then:
$$ P = \frac{3 \text{ trillion} \times 5}{10 \text{ trillion}} = 1.5 $$
- This means that the price level has increased from 1 to 1.5, indicating inflation.

---

# Quantity Theory of Money - example (cont.)

- Example: If the money supply (M) increases by 10% and velocity (V) is constant, then:
$$ \%M + \%V = \%P + \%Y $$
- If velocity (V) is constant, then:
$$ \%P = \%M - \%Y $$
- If output (Y) is constant, then:
$$ \%P = \%M $$
- If the money supply (M) increases by 10% and output (Y) is constant, then:
$$ \%P = 10\% $$
- If the money supply (M) increases by 10% and output (Y) increases by 5%, then:
$$ \%P = 10\% - 5\% = 5\% $$

- If the money supply (M) increases by 10% and output (Y) decreases by 5%, then:
$$ \%P = 10\% + 5\% = 15\% $$


---

# Determinants of Velocity

- Velocity (V) is influenced by several factors:
  - Payment technologies (e.g., credit cards, digital payments)
  - Financial innovations (e.g., money market funds)
  - Economic conditions (e.g., inflation, interest rates)

- Changes in velocity can affect the relationship between money supply and price level.

- For example, if velocity increases due to technological advancements, the same money supply can lead to higher output and lower prices.
- Conversely, if velocity decreases due to economic uncertainty, the same money supply may lead to higher prices and lower output.
- The determinants of velocity are important for understanding how changes in the money supply can impact the overall economy.

---

# Demand for Money
- The demand for money is the desire to hold cash or liquid assets for transactions, precautionary, and speculative purposes.
- The demand for money is influenced by several factors:
  - Interest rates
  - Income levels
  - Inflation expectations
  - Economic conditions

- The demand for money can be represented by the following equation:
$$ M_d = f(1/V, Y, P) $$
- Where:
  - $M_d$ = demand for money
  - $V$ = velocity of money
  - $Y$ = income level
  - $P$ = price level

---

class: my-one-page-font

# Fisher Equation and Money Demand

- Fisher’s interpretation:
  - Velocity is constant, so:

$$ P \times Y = M \times V $$

- Dividing by V:

$$ M = \frac{P \times Y}{V} $$

- Money demand is proportional to nominal income (PY).
- Implications for monetary policy:
  - If the central bank increases the money supply, it can lead to higher prices or output.
  - The relationship between money supply and price level is crucial for understanding inflation.

---

# Keynesian Theories of Money Demand

- Three motives for holding money:
  - Transactions Motive
  - Precautionary Motive
  - Speculative Motive

**Formula for Money Demand:**

$$ M_d = f(i, Y) $$

- Demand for money is negatively related to interest rates and positively related to income.
- Empirical evidence supports the liquidity preference theory, showing that money demand is sensitive to interest rates and income levels.
- The liquidity preference theory is a Keynesian approach to money demand that emphasizes the role of interest rates in determining the demand for money.
- It suggests that individuals prefer to hold money for transactions and precautionary purposes, but they also consider the opportunity cost of holding money in terms of foregone interest.
- The theory posits that the demand for money is inversely related to interest rates, meaning that as interest rates rise, the demand for money decreases.
- The liquidity preference theory is important for understanding how changes in interest rates can impact the overall economy.

---

# Inflation and Quantity Theory

- Rewriting the quantity equation in percentage change form:

$$ \%M + \%V = \%P + \%Y $$

- If velocity is constant:

$$ \%P = \%M - \%Y $$

- Implications for monetary policy and inflation targeting.
- If the money supply grows faster than output, it can lead to inflation.
- If the money supply grows slower than output, it can lead to deflation.
- The quantity theory of money is a classical economic theory that explains the relationship between the money supply and price levels in an economy.

---

# Budget Deficits and Inflation

- Financing government deficits:
  - Taxation
  - Borrowing
  - Money Creation (Seigniorage)

- Hyperinflation example: Zimbabwe (2000s).
- Hyperinflation occurs when there is excessive money supply growth without corresponding economic growth.
- It leads to a rapid increase in prices, eroding the purchasing power of money.
- Hyperinflation can result from various factors, including excessive government spending, loss of confidence in the currency, and external shocks.
- The consequences of hyperinflation include:
  - Decreased savings and investment
  - Increased uncertainty and volatility in the economy
  - Social and political instability

---

# Portfolio Theories of Money Demand

- Factors influencing money demand:
  - Wealth
  - Risk
  - Liquidity of other assets

- Empirical Evidence: Sensitivity of money demand to interest rates and income levels.
- Portfolio theories of money demand emphasize the role of risk and return in determining the demand for money.
- They suggest that individuals hold money as part of a diversified portfolio of assets.
- The demand for money is influenced by factors such as wealth, risk preferences, and the liquidity of other assets.
- Empirical evidence supports the idea that money demand is sensitive to interest rates and income levels.
- The portfolio theories of money demand provide a more comprehensive understanding of the factors that influence individuals' decisions to hold money.
- They highlight the importance of considering the trade-offs between holding money and other assets in a portfolio context.
- The portfolio theories of money demand are important for understanding how changes in interest rates and economic conditions can impact the overall demand for money in an economy.

---

# Summary and Review Questions

- Review the key concepts and equations.

- Application: How does the quantity theory of money help in understanding hyperinflation?




---

class: inverse, center, middle

# 2. Class Activity

---

class: inverse, center, middle

# Any QUESTIONS?

## Thank You!  

---

# Next Class

- (May 23) 
  - **Chap 23.** Aggregate Demand and Supply Analysis 


???
1. To print pdf slides
https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

pagedown::chrome_print("W1_ME.html") # but not all pictures are visible

2. Option: https://stackoverflow.com/questions/54968311/xaringan-export-slides-to-pdf-while-preserving-formatting

install.packages("remotes")
remotes::install_github("jhelvy/xaringanBuilder")
remotes::install_github("jhelvy/renderthis@v0.0.9")

library(xaringanBuilder)
build_pdf("DVC.html")

3. Option
writeBin(as.raw(c()), "favicon.ico") # create an empty favicon.ico file
install.packages("renderthis")
remotes::install_github('rstudio/chromote')
library(renderthis)

renderthis::to_pdf("W11_FIS.html")

getwd()
setwd("C:/Users/Iegor/OneDrive - kdis.ac.kr/Documents/GitHub/Sogang/2025/Spring/Financial Institutions and System/Week 11")
